---
title: "PFOMT Xtals"
author: "Benjamin Weigel"
date: "12.06.2015"
output:
html_document:
highlight: pygments
theme: flatly
---

A simple script to process the unit cell parameters and space groups captured by the rigacu software from the `*.log` files.
The actual data was extracted via `$ grep()`in this fashion:
`grep "Spacegroup\ number" -r --include \*.log ../Stuff/datasets/ > ~/IPB/thesis/usr/statistics/data/spacegroup.txt`

Used `grep "^[\ ]\{9\}Mosaicity:" -r --include \*.log ../Stuff/datasets/ > ~/IPB/thesis/usr/statistics/data/unitcellmosaicity.txt`
for the mosaicity (first occurence in *.log file from XDS)

From Rigaku software:
```{r}
library(magrittr)
library(stringr)
library(dplyr)

## process spacegroup
tmp <- readLines("../data/spacegroup.txt") %>% str_extract_all(pattern = "((dtrefine){1}[_]*[[:digit:]]*)|([/][[:digit:]]+([_][[:alpha:]]+)+[_]*[/])|[[:digit:]]+$") %>%
  lapply(., function(x){if(length(x)>2) return(x) else return(NULL)})

id <- tmp %>%
  lapply(X = ., FUN = "[", 1) %>% unlist %>% str_replace_all(pattern = "[/]|(_/)", replacement = "")

xtaldf <- data.frame(cbind(id, tmp %>%
                             lapply(X = ., FUN = "[", 2) %>% unlist))

xtaldf <- data.frame(cbind(xtaldf, tmp %>%
                             lapply(X = ., FUN = "[", 3) %>% unlist %>% as.numeric))

## process unit cell size
tmp <- readLines("../data/unitcelllen.txt") %>% str_extract_all(pattern = "((dtrefine){1}[_]*[[:digit:]]*)|([/][[:digit:]]+([_][[:alpha:]]+)+[_]*[/])|([[:digit:]]+[.][[:digit:]]+)")  %>%
  lapply(., function(x){if(length(x)>4) return(x) else return(NULL)})

xtaldf <- data.frame(cbind(xtaldf, tmp %>%
                             lapply(X = ., FUN = "[", 3:5) %>% unlist %>% as.numeric %>% matrix(ncol = 3)))

## process angles
tmp <- readLines("../data/unitcellangle.txt") %>% str_extract_all(pattern = "((dtrefine){1}[_]*[[:digit:]]*)|([/][[:digit:]]+([_][[:alpha:]]+)+[_]*[/])|([[:digit:]]+[.][[:digit:]]+)")  %>%
  lapply(., function(x){if(length(x)>4) return(x) else return(NULL)})

xtaldf <- data.frame(cbind(xtaldf, tmp %>%
                             lapply(X = ., FUN = "[", 3:5) %>% unlist %>% as.numeric %>% matrix(ncol = 3)))

## process angles
tmp <- readLines("../data/unitcellmosaicity.txt") %>% str_extract_all(pattern = "((dtrefine){1}[_]*[[:digit:]]*)|([/][[:digit:]]+([_][[:alpha:]]+)+[_]*[/])|([[:digit:]]+[.][[:digit:]]+)")  %>%
  lapply(., function(x){if(length(x)>2) return(x) else return(NULL)})

xtaldf <- data.frame(cbind(xtaldf, tmp %>%
                             lapply(X = ., FUN = "[", 3) %>% unlist %>% as.numeric %>% matrix(ncol = 1)))

names(xtaldf)[-1] <- c("dtref", "spacegroup.num", "len.a", "len.b", "len.c", "ang.a", "ang.b", "ang.c", "mosaicity")

xtaldf %>% mutate(volume = len.a*len.b*len.c)
head(xtaldf)
```

From Refmac:

`grep "\(Cell\ from\ mtz\)\|\(Space\ group\ from\ mtz\)" -r --include \*refmac5.log > ~/IPB/thesis/usr/statistics/data/refmac.txt`

```{r}
refmacdf <- readLines("../data/refmac.txt") %>% 
  str_extract_all(pattern = "(^([[:alnum:]]+)([_]*[[:alnum:]]+)+([.]1)?[_]?[/])|([[:digit:]]+[.][[:digit:]]+)|([-][ ]+[[:digit:]]+)|([-][ ]+[[:alpha:]]{1}[ ]+([[:digit:]]+[ ]*)+)|([[:digit:]]+_refmac)") %>% matrix(ncol = 2, byrow = T) %>%
  apply(., 1, unlist) %>% t %>% data.frame(stringsAsFactors = F)

refmacdf %<>% select(-c(9,10)) %>%
  mutate(X1 = str_replace(X1,"[/]", ""),
         X2 = str_replace(X2,"[_]refmac", ""),
         X11 = str_replace(X11,"[-][ ]+", ""),
         X12 = str_replace(X12,"[-][ ]+", ""))

names(refmacdf) <- c("id", "refmac.run", "len.a", "len.b", "len.c", "ang.a", "ang.b", "ang.c", "spacegroup.num", "spacegroup.name")

refmacdf[2:8] <- transform(refmacdf[2:8], 
          len.a = as.numeric(len.a),
          len.b = as.numeric(len.b),
          len.c = as.numeric(len.c),
          ang.a = as.numeric(ang.a),
          ang.b = as.numeric(ang.a),
          ang.c = as.numeric(ang.a),
          refmac.run = as.numeric(refmac.run))

## to get the latest refmac run results
refmacdf$id <- factor(refmacdf$id, levels = unique(refmacdf$id), ordered = T)
refmacdf %<>% group_by(id) %>% mutate(max=which(refmac.run == max(refmac.run)))

tmp <- refmacdf %>% group_by(id) %>% summarize(n = n(), max=unique(max))
tmp <- c(0,cumsum(tmp$n)[-5]) + tmp$max

refmacdf <- refmacdf[tmp,]

print(refmacdf)
```

