---
title: "WEB272 -- SOMT Refolding"
author: "Benjamin Weigel"
date: "07/23/2015"
output:
  pdf_document:
    number_sections: yes
    toc: yes
---

After rebuffering protein concentrations were measured by Bradford:

```{r, cache=FALSE, echo=FALSE}
knitr::opts_chunk$set(dev='tikz', error=T, warning=F, message=F, size='tiny', echo=F)
```

# Protein concetration #
```{r, fig.height=3}
library(magrittr)
library(dplyr)
library(ggplot2)

load("/media/mori/IPB/WEB272/140625_ResponseData.Rda")
a <- 0.0055
b <- 0.7114

df <- SOMT2_140625[,1:20]
df$BFmean <- c(14.81, -4.02, -2.26, 30.84, 62.27, 26.47, 66.65, 4.03, -2.93, -5.45, -6.27, 10.71)
df$BFsd <- c(3.48, 3.91, 4.25, 3.22, 1.20, 5.02, 5.58, 5.07, 0.73, 2.06, 1.16, 0.60)

df$BFmean[which(df$BFmean <= 0)] <- 0
df$BFsd[which(df$BFmean <= 0)] <- 0

df$sample <- as.factor(1:12)

ggplot(df, aes(x=sample, y=BFmean)) +   
  geom_bar(stat="identity", fill="coral") +
  geom_errorbar(aes(ymin=BFmean-BFsd, ymax=BFmean+BFsd), width=0) +
  labs(x="buffer", y="protein concentration $(\\frac{\\mu g}{ml})$") +
  theme_bw()
```

## Regression tree ##
Next we build a regression tree. To see, which factors have the most influence on the protein concentration.

```{r, dev='CairoPDF', echo=c(4,5,6)}
library(rpart)
library(rpart.plot)
library(rattle)

somttree2 <- rpart(BFmean ~ ., 
                data=df[,c(1:11,21)],
                control = c(minsplit = 3,
                            minbucket = 2,
                            complexity = 0.001))
fancyRpartPlot(somttree2, digits=3)
```

Arginine seems to have the biggest impact on refolding efficiency (Arginine addition is better). Then comes SAH (no SAH is better).

## Main effects plot ##

```{r, dev='CairoPDF'}
library(FrF2)
SOMT2_140625$BFmean <- df$BFmean

RES <- response.names(SOMT2_140625)
response.names(SOMT2_140625) <- c(RES, "BFmean")

ME <- MEPlot(SOMT2_140625, abbrev=6, select=c(1:8), response="BFmean")

MEgg <- tidyr::gather(cbind(as.data.frame(ME), state = rownames(ME)), key = ME, value = value, -state)
```

```{r, echo=F, fig.height=3}
ggplot(MEgg, aes(x=state, y=value)) + 
  theme_bw() +
  geom_line(aes(group = ME), color="cornflowerblue", size=1) +
  geom_point(size=3) + 
  facet_wrap(~ME, nrow=1) +
  labs(y="mean protein concentration $(\\frac{\\mu g}{ml})$", x="factor level") +
  geom_hline(aes(yintercept=mean(MEgg$value)), linetype=2, size=1)
```

## Statistical test ##

Test the statistical significance of main effects.
Only Arginine is statistically significant to a p-value of 0.05. SAH to a p-value of 0.1.

```{r, results='asis'}
test.lm <- lm(BFmean ~ Arginine + pH + Glycerin + ionicStrength + divCations + redox + CycloDex + SAH, data=df[,c(1:11,21)])
aov(test.lm) %>% summary %>% xtable::xtable()

test.lm <- lm(BFmean ~ Arginine + pH + Glycerin + ionicStrength + divCations + redox + CycloDex + SAH + e1 +e2 + e3, data=df[,c(1:11,21)])
LenthPlot(test.lm, alpha=0.05)
```

# Protein volume activity #
```{r, fig.height=3}
load("/media/mori/IPB/WEB272/HPLC.Rda")

df <- SOMT2_140625[,c(1:11, 17)]
df$sample <- as.factor(1:12)

ggplot(df, aes(x=sample, y=P1_AC)) +   
  geom_bar(stat="identity", fill="coral") +
  labs(x="buffer", y="$\\frac{\\mathrm{ponciretin}}{\\mathrm{AC-9}} (\\mathrm{AU})$") +
  theme_bw()

ggplot(df, aes(x=sample, y=(INTDATA$Product_1 / INTDATA$'AC-9') / (INTDATA$Naringenin / INTDATA$'AC-9'))) +   
  geom_bar(stat="identity", fill="coral") +
  labs(x="buffer", y="$\\frac{\\mathrm{ponciretin}}{\\mathrm{AC-9}} (\\mathrm{AU})$") +
  theme_bw()
```

## Regression tree ##
Next we build a regression tree. To see, which factors have the most influence the SOMT activity.

```{r, dev='CairoPDF', echo=c(4,5,6)}
library(rpart)
library(rpart.plot)
library(rattle)

somttree2 <- rpart(P1_AC ~ ., 
                data=df[,c(1:8,12)],
                control = c(minsplit = 3,
                            minbucket = 2,
                            complexity = 0.001))
fancyRpartPlot(somttree2, digits=3)
```

The redox status seems to have the biggest impact on refolding efficiency measured by activity (reducing is better, DTT). Then comes arginine (arginine is better).

## Main effects plot ##

```{r, dev='CairoPDF'}
library(FrF2)
test.lm <- lm(P1_AC ~ pH + Arginine + Glycerin + ionicStrength + divCations + redox + CycloDex + SAH, data=df[,c(1:8,12)])
ME <- MEPlot(test.lm, abbrev=6, select=c(1:8), response="P1_AC")

MEgg <- tidyr::gather(cbind(as.data.frame(ME), state = rownames(ME)), key = ME, value = value, -state)
```

```{r, echo=F, fig.height=3}
ggplot(MEgg, aes(x=state, y=value)) + 
  theme_bw() +
  geom_line(aes(group = ME), color="cornflowerblue", size=1) +
  geom_point(size=3) + 
  facet_wrap(~ME, nrow=1) +
  labs(y="mean $\\frac{\\mathrm{ponciretin}}{\\mathrm{AC-9}} (\\mathrm{AU})$", x="factor level") +
  geom_hline(aes(yintercept=mean(MEgg$value)), linetype=2, size=1)

```

## Statistical test ##

Test the statistical significance of main effects.
Only Arginine is statistically significant to a p-value of 0.05. SAH to a p-value of 0.1.

```{r, results='asis'}
aov(test.lm) %>% summary %>% xtable::xtable()

test.lm <- lm(P1_AC ~ Arginine + pH + Glycerin + ionicStrength + divCations + redox + CycloDex + SAH + e1 +e2 + e3, data=df[,c(1:12)])
LenthPlot(test.lm, alpha=0.05)
```


